A template for the *arxiv* style

### 2021-12-16 main.Rmd
clean up

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  fig.path = "figures/",
  fig.height = 10
)
options(prompt = "R> ", continue = "+ ")
```

@@

**Motivation**

@@

**Existing packages**

@@

**Our new data structure for spatio-temporal data**
This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}. \pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension. With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble\pkg{. The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link]. \newline

**Section division**

@@

Spatio-temporal data usually come in various forms and Figure \ref{fig:cubble-diagram} shows four examples of this. No matter which form the data is in, these formats share some common components that characterise spatio-temporal data. A spatial identifier (`id` in the diagram) is the unique identifier of each site. The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded. Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map. These identifiers will be the building blocks for the data structure introduced below. Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e. the name or code of the weather station and temporal variables that varies with time.

```{r cubble-diagram, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data. (1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
```

In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:def-cubble} sketches the two forms along with the associated attributes. The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier. The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e. the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form). Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group. This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.

@@

```{r def-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
```

@@

After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained. Figure \ref{fig:hierachical} illustrate the relationship between the long and nested form cubble at both site and cluster level. More description on this. Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute. `switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`. `stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.

@@

@@

@@

The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling]. Diagram \ref{fig:interactive-workflow} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble. When a user action is captured from the map, the site will be activated in the nested cubble. Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`. The long cubble will then highlight the  activated series in the time series plot.

The linking is also available from the time series plot to the map. The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble. All the observations that share the same `id`,  either in the long and nested cubble, are then activated. This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble. These activated observations will then being reflected in the updated plots and Diagram \ref{fig:interactive-workflow-2} in the Appendix illustrates this process.

```{r interactive-workflow, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
```

@@

Bureau of Meteorology (BoM) collects climate variables. This variables can be openly queried via `rnoaa`. `weatherdata` provides a readily available copy of several data.`weatherdata::historical_tmax` provides historical maximum temperature recorded for Australian stations with the earliest dating back to 1859 and this can be used to understand the change of maximum temperature from a past period to nowadays. Figure \ref{fig:glyphmap-workflow} shows the steps involved from the initial data to the final glyph map in Figure \ref{fig:glyphmap}.

```{r glyphmap-workflow, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon code"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
```

@@

```{r echo = FALSE, eval = FALSE}
tmax_nested <- weatherdata::historical_tmax %>%
  filter(between(stringr::str_sub(id, 7, 8), 46, 90))

tmax_long <- tmax_nested %>%
  stretch() %>%
  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
  mutate(
    month = lubridate::month(date),
    group = as.factor(ifelse(
      lubridate::year(date) > 2015,
      "2016 ~ 2020",
      "1971 ~ 1975"
    ))
  ) %>%
  group_by(month, group) %>%
  summarise(tmax = mean(tmax, na.rm = TRUE))

tmax_nested2 <- tmax_long %>%
  tamp() %>%
  filter(nrow(ts) == 24)

tmax_long2 <- tmax_nested2 %>%
  stretch() %>%
  migrate(latitude, longitude)

gly <- glyphs(tmax_long2, "longitude", "month", "latitude", "tmax",
              height = 0.6, width = 1)

nsw_vic <- ozmaps::abs_ste %>%
  filter(NAME %in% c("Victoria", "New South Wales")) #

plot_map(nsw_vic, fill = "transparent") +
  geom_path(data = gly, 
            aes(gx, gy, group = interaction(id, group), 
                color = group)) +
  scale_color_brewer(palette = "Dark2") + 
  theme_bw() + 
  coord_sf(xlim = c(141, 154)) + 
  theme(legend.position = "bottom")

ggsave(filename = "figures/basic-manipulation.png", width = 10, height = 10)
```

```{r glyphmap, out.width="100%", out.height="70%", fig.cap="asdfasd"}
knitr::include_graphics("figures/basic-manipulation.png")
```

@@

In figure \ref{fig:glyphmap}, there are a few stations around Sydney (151E, 34S) and New castle (152E, 33S). Aggregating them into averages or other statistics would give a better picture. This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.

In Figure \ref{fig:aggregation}

@@

```{r aggregation, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
knitr::include_graphics("figures/aggregation.png")
```

@@

```{r echo = FALSE}
river <- weatherdata::water %>%
  stretch() %>%
  select(date, Water_course_level) %>%
  tamp() %>%
  mutate(type = "river")

climate <- weatherdata::climate_full %>%
  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
  stretch() %>%
  filter(lubridate::year(date) == 2020) %>%
  tamp() %>%
  mutate(type = "climate")

vic_map <- rmapshaper::ms_simplify(ozmaps::abs_ste %>% filter(NAME == "Victoria"))
plot_map(vic_map) +
  geom_point(
    data = dplyr::bind_rows(river, climate),
    aes(x = long, y = lat, color = type)
  ) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw()
```

```{r echo = FALSE, fig.height=3, fig.cap = "shtishisaasdf"}
set.seed(123)
dt <- tibble(
  id = factor(c(rep("A", 31), rep("a", 31)), levels = c("A", "a")),
  date = rep(1:31, 2),
  val = c(
    c(rnorm(5), 10, rnorm(7), 5, rnorm(8), 7, rnorm(8)),
    c(
      rnorm(6, sd = 0.5), 7, rnorm(7, sd = 0.5), rnorm(6, sd = 0.5),
      4, rnorm(5, sd = 0.5), 6, rnorm(4, sd = 0.5)
    )
  )
) %>% mutate(val = ifelse(val < 0, 0, val))

circle <- tibble(
  x = c(6, 14, 23, 7, 21, 27),
  y = c(10, 5, 7, 7, 4, 6),
  id = factor(c(rep("A", 3), rep("a", 3)), levels = c("A", "a")),
  xend = x + 5
)

errorbar <- bind_rows(
  circle %>% filter(id == "A"),
  circle %>% filter(id == "A") %>% mutate(id = "a")
) %>%
  mutate(id = factor(id, c("A", "a")))

fallin <- tibble(
  id = factor(rep("a", 2), c("A", "a")),
  x = c(7, 27),
  y = -0.5,
  label = "\U2714"
)

ggplot() +
  geom_line(data = dt, aes(x = date, y = val, group = id), color = "grey60") +
  geom_vline(data = circle, aes(xintercept = x), linetype = "longdash", color = "grey60") +
  geom_errorbar(data = errorbar, aes(xmin = x, xmax = xend, y = -0.5), width = 0.5) +
  geom_point(data = circle, aes(x = x, y = y), color = "brown") +
  geom_text(data = fallin, aes(x = x, y = y, label = label), color = "brown", size = 5) +
  facet_wrap(vars(id), nrow = 2) +
  scale_y_continuous(breaks = seq(0, 10, 2)) +
  scale_x_continuous(breaks = seq(1, 31, 1)) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  labs(x = "Time", y = "Value")
```

```{r}
river <- river %>%
  stretch() %>%
  rename(prcp = Water_course_level) %>%
  tamp()
```

Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`. We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`. We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`. This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points). `temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.

```{r}
res <- match_sites(river, climate,
  temporal_var_to_match = prcp,
  temporal_independent = climate,  
  temporal_n_highest = 30,
  temporal_min_match = 15
)

res
```

@@

```{r echo = FALSE, out.width="100%", fig.cap="..."}
p1 <- plot_map(vic_map) +
  geom_point(
    data = res,
    aes(x = long, y = lat, color = type)
  ) +
  ggrepel::geom_label_repel(
    data = res %>% filter(type == "river"),
    aes(x = long, y = lat, label = .group)
  ) +
  scale_color_brewer(palette = "Dark2") +
  ggtitle("Victoria") +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

res_long <- res %>%
  stretch(ts) %>%
  migrate(.group, type) %>%
  mutate(prcp = scale(prcp)[, 1]) %>%
  switch_key(.group) %>%
  migrate(type)

p2 <- res_long %>%
  ggplot(aes(
    x = date, y = prcp,
    group = id, color = type
  )) +
  geom_line() +
  facet_wrap(vars(.group)) +
  scale_color_brewer(palette = "Dark2", guide = "none") +
  theme_bw() +
  labs(x = "week") +
  scale_x_date(date_labels = "%b")

p1 | p2
```

@@

```{r eval = FALSE}
nc <- ncdf4::nc_open(here::here("data/adaptor.mars.internal-1638344933.932894-4194-11-a5d35ac3-b36d-4982-afa5-a5dd193a6718.nc"))

nc <- ncdf4::nc_open(here::here("data/tas_Amon_ACCESS-CM2_ssp585_r1i1p1f1_gn_20150116-20201216_v20191108.nc"))
temp <- as_cubble(nc, vars = "tas")
vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")

res <- temp %>% 
  stretch(ts) %>% 
  mutate(d2m = d2m - 273.15) %>% 
  filter(lubridate::hour(time) == 12) %>% 
  group_by(id) %>% 
  summarise(d2m = mean(d2m, na.rm= TRUE)) %>%
  migrate(long, lat)

vic_string <- vic_map %>% sf::st_cast("MULTILINESTRING")

res %>% 
  ggplot() +
  geom_tile(aes(x = long, y = lat, fill = d2m)) +
  geom_sf(data = vic_string , alpha = 0.5, fill = "transparent") +
  theme_bw() + 
  colorspace::scale_fill_continuous_sequential("Mint")

```

@@

```{r echo = FALSE, out.width="100%", out.height="30%"}
knitr::include_graphics(here::here("figures/linking.png"))
```

```{r echo = FALSE, out.width="50%", out.height="30%"}
knitr::include_graphics(here::here("figures/popup.png"))
```

@@

```{r interactive-workflow-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.007.png"))
```

@@

# Old stuff

In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [... ] workflow to work with temporal data. In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier. However, a tsibble object, by construction, always requires the `index` in its structure. This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e. features of each series) don't have the time dimension. Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects. In addition, the long form structure of a tsibble object means spatio variables (i.e. longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp. This repetition is unnecessary and would inflate the object size for long series.

# Create a cubble

The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier. `climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier. To create a cubble for this data, use:

```{r}
climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat))
```

Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata. \textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata. Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list. As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne. We first create a station data frame with the 5 target stations.

```{r echo = FALSE}
latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631)

cand <- aus_climate %>%
  as_tibble() %>%
  select(id, lat, long, elev, name, wmo_id)

calc_dist <- function(df_ref, df, n = 5) {
  long_ref <- df_ref$long
  lat_ref <- df_ref$lat

  df %>%
    mutate(
      long_ref = long_ref,
      lat_ref = lat_ref,
      dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref)
    ) %>%
    slice_min(dist, n = n) %>%
    mutate(city = df_ref$id) %>%
    select(-long_ref, -lat_ref)
}

(stations <- do.call("rbind", purrr::map(
  1:nrow(latlon_ref),
  ~ calc_dist(latlon_ref[.x, ], cand)
)))
```

We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example.

```{r eval = FALSE}
sydmel_climate <- stations %>%
  rowwise() %>%
  mutate(ts = list(meteo_pull_monitors(id,
    date_min = "2020-01-01",
    date_max = "2020-12-31",
    var = c("PRCP", "TMAX", "TMIN")
  ) %>%
    select(-id))) %>%
  as_cubble(key = id, index = date, coords = c(long, lat))
```

```{r echo = FALSE}
(sydmel_climate <- stations %>%
  left_join(aus_climate) %>%
  as_cubble(key = id, index = date, coords = c(long, lat)))
```

Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020. Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively.

With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way. The table below shows how some `dplyr` verbs are mapped into the operation in a cubble. With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`. [talk about why it is useful]

## Cubble operations

```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
```

>>>>

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  fig.path = "figures/",
  fig.height = 10
)
options(prompt = "R> ", continue = "+ ", 
        tibble.print_max = 5, tibble.print_min = 5)
```

@@

@@

@@

This paper describes the implementation of a new spatio-temporal data structure: \pkg{cubble}. \pkg{cubble} implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension. With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to \pkg{cubble}. The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link]. \newline

@@

Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this. No matter which form the data is in, these formats share some common components that characterise spatio-temporal data. A spatial identifier (`id` in the diagram) is the unique identifier of each site. The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded. Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map. These identifiers will be the building blocks for the data structure introduced below. Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e. the name or code of the weather station and temporal variables that varies with time.

```{r illu-input, echo = FALSE, out.width = "100%", fig.cap="Illustration of incoming data formats for spatio-temporal data. (1) Data comes in as a single table; (2) Separate tables for spatial and temporal variables; (3) A single table with all the parameters used to query the database and a separate table for queried data; and (4) Cubical data in array or NetCDF format.", fig.align='center', out.height = "15%"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.001.png"))
```

In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes. The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier. The nested cubble is suitable for working with operations that are only identified by site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e. the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form). Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group. This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.

@@

```{r illu-cubble, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Illustration of nested and long cubble."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
```

@@

After we have got `cluster_nested`, spatial and temporal data at both levels can be easily obtained. Figure \ref{fig:illu-hier} illustrate the relationship between the long and nested form cubble at both site and cluster level. More description on this. Start with the original `station_nested`, `stretch()` expands the `ts` column with each station (`id`) forming a group and attach variables invariant to `id` as an attribute. `switch_key()` changes the `key` from `id` to `cluster` and nests all the spatial variables that variant to `cluster`. `stretch()` `cluster_nested` will store variables that are invariant to `cluster` as a tibble in the attribute.

@@

```{r illu-hier, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "Hierarchical structure"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
```

@@

Figure \ref{fig:illu-matching}

```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
knitr::include_graphics("figures/illu-matching.png")
```

@@

The cubble structure fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling]. Diagram \ref{fig:illu-interactive} illustrates how linking works with the two forms in a cubble, where a time series plot is created with the long cubble and a map is created with the nested cubble. When a user action is captured from the map, the site will be activated in the nested cubble. Then, the nested cubble will communicate to the long cubble to activate all the observations with the same `id`. The long cubble will then highlight the  activated series in the time series plot.

The linking is also available from the time series plot to the map. The selection on the time series is through selecting the point on the time series and once a point is selected, it will be activated in the long cubble. All the observations that share the same `id`,  either in the long and nested cubble, are then activated. This includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble. These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.

```{r illu-interactive, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
```

@@

Bureau of Meteorology (BoM) collects climate variables. This variables can be openly queried via `rnoaa`. `weatherdata` provides a readily available copy of several data. `weatherdata::historical_tmax` provides historical maximum temperature recorded for Australian stations with the earliest dating back to 1859 and this can be used to understand the change of maximum temperature from a past period to nowadays.

Figure \ref{fig:basic-manip}.

```{r eval=  FALSE, echo = TRUE}
tmax <- weatherdata::historical_tmax %>%
  # filter out VIC and NSW stations
  filter(between(stringr::str_sub(id, 7, 8), 46, 90)) %>%

  # filter out observations in between 1971-1975 and 2016 - 2020
  stretch() %>%
  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) %>%
  mutate(month = lubridate::month(date), 
         group = as.factor(ifelse(lubridate::year(date) > 2015, 
                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
  group_by(month, group) %>%
  # summarise tmax into monthly mean for both periods
  summarise(tmax = mean(tmax, na.rm = TRUE)) %>%

  # filter out stations that have complete records in both periods 
  tamp() %>%
  filter(nrow(ts) == 24) %>%

  # migrate longitude and latitude to the long form for making the glyph map
  stretch() %>%
  migrate(latitude, longitude)

gly <- glyphs(tmax, "longitude", "month", "latitude", "tmax",
              height = 0.6, width = 1)

nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))

ggplot() + 
  geom_sf(data = nsw_vic, fill = "transparent") + 
  geom_path(data = gly, 
            aes(gx, gy, group = interaction(id, group), color = group)) +
  scale_color_brewer(palette = "Dark2") + 
  theme_bw() + 
  coord_sf(xlim = c(140, 155)) + 
  theme(legend.position = "bottom")
```

```{r basic-manip, out.width="100%", out.height="70%", fig.cap="asdfasd"}
knitr::include_graphics("figures/basic-manip.png")
```

@@

@@

In figure \ref{fig:basic-manip}, there are a few stations around Sydney (151E, 34S) and New castle (152E, 33S). Aggregating them into averages or other statistics would give a better picture. This can be done with cubble through a hierarchical structure, where the cluster indicator is one level above the individual station.

In Figure \ref{fig:basic-agg}

@@

```{r basic-agg, out.width="100%", out.height="90%", fig.cap="sdfasdf"}
knitr::include_graphics("figures/basic-agg.png")
```

@@

Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`. We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`. We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`. This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points). `temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``

@@

```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
knitr::include_graphics("figures/matching.png")
```

@@

@@

Figure \ref{fig:interactive-linking} 
Figure \ref{fig:interactive-popup}

```{r interactive-linking, echo = FALSE, out.width="100%", out.height="30%", fig.cap = "sdfasd"}
knitr::include_graphics(here::here("figures/linking.png"))
```

```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "sdfasd"}
knitr::include_graphics(here::here("figures/popup.png"))
```

@@

```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "demon interactivity"}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
```

@@

<!-- # Old stuff -->

<!-- In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [... ] workflow to work with temporal data. In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier. However, a tsibble object, by construction, always requires the `index` in its structure. This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e. features of each series) don't have the time dimension. Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects. In addition, the long form structure of a tsibble object means spatio variables (i.e. longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp. This repetition is unnecessary and would inflate the object size for long series. -->

<!-- # Create a cubble -->

<!-- The creation of a cubble requires the site identifier (`key`), as well as the spatial (`coords`) and temporal (`index`) identifier. `climate_flat` is already a tibble and it uses `id` to identify each station, `date` as the time identifier, and `c(long, lat)` as the spatial identifier. To create a cubble for this data, use: -->

<!-- ```{r} -->
<!-- climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat)) -->
<!-- ``` -->

<!-- Most of the time, spatio-temporal data doesn't come into this form and analysts need to query the climate variables based on station metadata. \textcolor{red}{This is also a problem illustrated in Section 3.5 in @tidydata. Here we provide a structured way to query this data based on the row-wise operator and nested list.} For this type of task, one can structure a metadata into a tibble and use row-wise operator to query the climate variables into a nested list. As an example here we demonstrate the workflow to find the 5 closest stations to Melbourne. We first create a station data frame with the 5 target stations. -->

<!-- ```{r echo = FALSE} -->
<!-- latlon_ref <- tibble::tibble(id = "melbourne", lat = -37.8136, long = 144.9631) -->

<!-- cand <- aus_climate %>% -->
<!--   as_tibble() %>% -->
<!--   select(id, lat, long, elev, name, wmo_id) -->

<!-- calc_dist <- function(df_ref, df, n = 5) { -->
<!--   long_ref <- df_ref$long -->
<!--   lat_ref <- df_ref$lat -->

<!--   df %>% -->
<!--     mutate( -->
<!--       long_ref = long_ref, -->
<!--       lat_ref = lat_ref, -->
<!--       dist = rnoaa::meteo_spherical_distance(lat, long, lat_ref, long_ref) -->
<!--     ) %>% -->
<!--     slice_min(dist, n = n) %>% -->
<!--     mutate(city = df_ref$id) %>% -->
<!--     select(-long_ref, -lat_ref) -->
<!-- } -->

<!-- (stations <- do.call("rbind", purrr::map( -->
<!--   1:nrow(latlon_ref), -->
<!--   ~ calc_dist(latlon_ref[.x, ], cand) -->
<!-- ))) -->
<!-- ``` -->

<!-- We can query the climate information into a nested list named `ts` for each station with the `rowwise()` operator. To create a cubble, supply the same identifiers as with the first example. -->

<!-- ```{r eval = FALSE} -->
<!-- sydmel_climate <- stations %>% -->
<!--   rowwise() %>% -->
<!--   mutate(ts = list(meteo_pull_monitors(id, -->
<!--     date_min = "2020-01-01", -->
<!--     date_max = "2020-12-31", -->
<!--     var = c("PRCP", "TMAX", "TMIN") -->
<!--   ) %>% -->
<!--     select(-id))) %>% -->
<!--   as_cubble(key = id, index = date, coords = c(long, lat)) -->
<!-- ``` -->

<!-- ```{r echo = FALSE} -->
<!-- (sydmel_climate <- stations %>% -->
<!--   left_join(aus_climate) %>% -->
<!--   as_cubble(key = id, index = date, coords = c(long, lat))) -->
<!-- ``` -->

<!-- Below are the how the nested and long form look like for Australia climate data, which records daily precipitation, maximum and minimum temperature for 55 stations across Australia from 2015- 2020. Notice that each station forms a group in both forms and specifically, the nested and long form have a underlying `rowwise_df` and `grouped_df` respectively. -->

<!-- With a cubic framework on mind, different types of manipulation with cubble can be thought of as slicing the cube in various way. The table below shows how some `dplyr` verbs are mapped into the operation in a cubble. With the existing grouping on the station, additional grouping can be added with `group_by` and removed with `ungrouped`. [talk about why it is useful] -->

<!-- \newpage -->

<!-- ## Cubble operations -->

<!-- ```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'} -->
<!-- knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png")) -->
<!-- ``` -->
